---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dcmodifydb

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/dcmodifydb)](https://CRAN.R-project.org/package=dcmodifydb)
[![R-CMD-check](https://github.com/data-cleaning/dcmodifydb/workflows/R-CMD-check/badge.svg)](https://github.com/data-cleaning/dcmodifydb/actions)
[![Codecov test coverage](https://codecov.io/gh/data-cleaning/dcmodifydb/branch/main/graph/badge.svg)](https://codecov.io/gh/data-cleaning/dcmodifydb?branch=main)
<!-- badges: end -->

The goal of dcmodifydb is to apply modification rules specified with `dcmodify` 
on a database table, allowing for documented, reproducable data cleaning adjustments
in a database.

## Installation

<!-- You can install the released version of dcmodifydb from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("dcmodifydb") -->
<!-- ``` -->

The development version from [GitHub](https://github.com/) can be installed with:

``` r
# install.packages("devtools")
devtools::install_github("data-cleaning/dcmodifydb")
```
## Example


```{r, code = readLines("./example/modify.R")}
```


### Documented rules


```{r}
library(DBI)
library(dcmodify)
library(dcmodifydb)
con <- dbConnect(RSQLite::SQLite())
```


You can use YAML to store the modification rules: "example.yml"

```yaml
rules:
- expr: if (age > 130) age <- 130
  name: M1
  label: 'Maximum age'
  description: |
    Human age is limited.
- expr: if (age < 12) {income <- 0}
  name: M2
  label: 'No Child Labor'
  description: |
    Children should not work.
```

```{r, eval = FALSE}
m <- modifier(.file = "example.yml")
```

```{r, echo = FALSE, eval = TRUE}
m <- modifier(.file = "example/example.yml")
```

```{r}
"age, income
  11,   2000
 150,    300
  25,   2000
" -> csv
income <- read.csv(text = csv, strip.white = TRUE)
dbWriteTable(con, "income", income)
tbl_income <- dplyr::tbl(con, "income")

tbl_income
modify(tbl_income, m, copy = FALSE)
```

Note: Modification rules can be written to yaml with `as_yaml` and `export_yaml`.

```{r, eval = FALSE}
dcmodify::export_yaml(m, "cleaning_steps.yml")
```

Generated sql can be written with `dump_sql`

```{r}
dump_sql(m, tbl_income)
```


```{r}
dbDisconnect(con)
```
