---
title: "dcmodifydb"
subtitle: "Documented reproducible data correction on a database"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dcmodifydb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The goal of dcmodifydb is to apply modification rules specified with `dcmodify` 
on a database table, allowing for documented, reproducable data cleaning adjustments
in a database.

`dcmodify` separates **intent** from **execution**: a user specifies the _what_
, _why_ and _how_ of an automatic data change and uses `dcmodifydb` to execute 
them on a `tbl` database table.

`dcmodidfydb` is optimized and restricted to database tables that can be accessed 
within R through `DBI`. It uses the `dbplyr` package to translate the 
data correction rules in R syntax into SQL syntax. The advantage of this 
approach is that all data correction is done within the database, which may be 
a requirement of your organisation or because the data table is simply too 
large to be held in memory. 
A disadvantage is that not all R statements can be translated into SQL statement,
so `dcmodifydb` is more restricted than `dcmodify` which can use the full R
potential. Nonetheless `dcmodifydb` may be sufficient and efficient for many
use cases.

For common error scenario's see `vignette("scenarios", package="dcmodifydb")`. 
For the supported syntax for specifying rules see `vignette("syntax", package="dcmodifydb")`.

## Installation

`dcmodifydb` can be installed with

```{r, eval=FALSE}
install.packages("dcmodifydb")
```
and loaded with:

```{r setup}
library(dcmodify)
library(dcmodifydb)
```

## Usage

`dcmodifydb` works on a database table, so we need a connection to a table within
a database. 

We set up a database table with sqlite using the `person` data set, but for your 
use case you should connect to your database.

```{r, echo=FALSE}
knitr::kable(person)
```


```{r}
con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
dplyr::copy_to(con, person)
```


We now retrieve a handle to the person table using `dplyr`

```{r}
person_tbl <- dplyr::tbl(con, "person")
person_tbl
```

The person dataset clearly contains some errors that can be corrected. 
We specify the corrections using modifier rules and apply them directly with 
the function`modify_so`.

First we correct that children can not have an income and that year must be a
long year.

```{r}
library(dcmodify) # needed for modifying rules
library(dcmodifydb) # needed to translate the rules
modify_so( person_tbl
         , if (age < 16) income = 0
         , if (year < 25) year = year + 2000
         )
```

Note that the corrections are made on a copy of the table by default, to avoid 
accidents with the data.

A better approach than directly applying corrections is to store the rules 
in a `modifier` object and apply them in a separate step to a data base table.

This makes it easier to maintain, use and document a set of rules.
With `dcmodify` one can specify rules with the function `modifier`:

```{r}
# separate rule set
m <- modifier( if (age < 16) income = 0
             , if (year < 25) year = year + 2000
             , if (cigarettes > 0 ) smokes = "yes"
             , if (smokes == "no") cigarettes = 0
             , ageclass <- if (age < 18) "child" else "adult"
             , gender <- switch( toupper(gender)
                               , "F" = "F"
                               , "V" = "F" # common mistake
                               , "M" = "M"
                               , "NB"
                               )
             )
```

`m` is now a set of rules that can be applied to a `data.frame` or `tbl`.

```{r}
print(m)
```

- M3 corrects `smokes` is "no" for persons who smoke cigarettes.

- M4 sets an unknown `cigarettes` to zero for non-smokers. 

- M5 shows an example of deriving a new variable (`ageclass`) from existing 
variables (`age`).

- M6 shows an example of recoding, changing labels of categories to the allowed
set of labels.

```{r}
# modify a copy of the table
modify(person_tbl, m, copy = TRUE)
```

Note that the rules are executed sequentially, in the order that they are gven.
For example the order of rule M3 and M4 matters: Rule M3 will change record 1 to a
smoker, while rule M4 would set the number of cigarettes to 0. 
This is intentional: correction rules often have an order in which they have to be
applied.

A nice properties of `modifier` rules, is that they can store extra metadata.
They have a `name`, `label` and `description` that can be used to describe the
intention and the why of a rule. An easy way of describing these properties is by
exporting the ruleset to yaml and specify the rules using the yaml file.

```{r, eval=FALSE}
export_yaml(m, "corrections.yml")
```

In the export yml file we can label and describe the rules, but also add new 
rules. Note that label and description are optional, but very much encouraged.

_corrections.yml_
```yaml
```{r, result="asis", child="../inst/db/corrections.yml"}
```
```

We can load these rules with:

```{r, eval = FALSE}
m <- modifier(.file = "corrections.yml")
modify(person_tbl, m, copy = TRUE)
```

```{r, echo = FALSE}
m <- modifier(.file = system.file("db/corrections.yml", package="dcmodifydb"))
modify(person_tbl, m, copy = TRUE)
```

`modify` translates the modification rules into SQL code and executes the sql 
queries on the database. For documentation or implementation purpose it can
be useful to see the generated sql code, with the documented rules. 

```{r dump, eval=FALSE}
dump_sql(m, person_tbl, file = "corrections.sql")
```

_corrections.sql_:
```SQL
```{r, eval=TRUE, echo=FALSE, results='asis'}
dump_sql(m, person_tbl)
```

```



