---
title: "dcmodifydb"
subtitle: "Documented reproducible data correction on a database"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dcmodifydb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The goal of dcmodifydb is to apply modification rules specified with `dcmodify` 
on a database table, allowing for documented, reproducable data cleaning adjustments
in a database.

`dcmodify` separates **intent** from **execution**: a user specifies the _what_
, _why_ and _how_ of an automatic data change and uses `dcmodifydb` to execute 
them on a `tbl` database table.

`dcmodidfydb` is optimized and restricted to database tables that can be accessed 
within R through `DBI`. It uses the `dbplyr` package to translate the 
data correction rules in R syntax into SQL syntax. The advantage of this 
approach is that all data correction is done within the database, which may be 
a requirement of your organisation or because the data table is simply too 
large to be held in memory. 
A disadvantage is that not all R statements can be translated into SQL statement,
so `dcmodifydb` is more restricted than `dcmodify` which can use the full R
potential. Nonetheless `dcmodifydb` may be sufficient and efficient for many
use cases.

For common error scenario's see `vignette("scenarios", package="dcmodifydb")`. 
For the supported syntax for specifying rules see `vignette("syntax", package="dcmodifydb")`.

## Installation

`dcmodifydb` can be installed with

```{r, eval=FALSE}
install.packages("dcmodifydb")
```
and loaded with:

```{r setup}
library(dcmodify)
library(dcmodifydb)
```

## Usage

`dcmodifydb` works on a database table, so we need a connection to a table within
a database. 

We set up a database table with sqlite using the person data set, but for your 
use case you should connect to your database.

```{r, echo=FALSE}
knitr::kable(person)
```


```{r}
con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
dplyr::copy_to(con, person)
```


We know retrieve a handle to the person table using `dplyr`

```{r}
person_tbl <- dplyr::tbl(con, "person")
person_tbl
```

The person dataset clearly contains some errors that should be corrected. 
We specify the corrections using modifier rules with the function `modifier`.

First we correct that minors should not have an income and that year should be a
long year.

```{r}
library(dcmodify) # needed for modifying rules
library(dcmodifydb) # needed to translate the rules
modify_so( person_tbl
         , if (age < 16) income = 0
         , if (year < 25) year = year + 2000
         )
```

Note that the corrections are made on a copy of the table by default, to avoid 
accidents with the data. 
Furthermore, directly applying the rules is not optimal. It is better to specifiy the
rules separately and document the rules.

```{r}
# separate rule set
m <- modifier( if (age < 16) income = 0
             , if (year < 25) year = year + 2000
             , if (cigarrettes > 0 ) smokes = "yes"
             , if (smokes == "no") cigarrettes = 0
             , ageclass <- if (age < 18) "child" else "adult"
             , gender <- toupper(gender)
             , gender <- switch( gender
                               , "V" = "F"
                               , gender
                               )
             )

# modify a copy of the table
modify(person_tbl, m, copy = TRUE)
```


```{r}
dump_sql(m, person_tbl)
```



