---
title: "dcmodifydb"
subtitle: "Documented reproducible data correction on a database"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dcmodifydb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The goal of dcmodifydb is to apply modification rules specified with `dcmodify` 
on a database table, allowing for documented, reproducable data cleaning adjustments
in a database.

`dcmodify` separates **intent** from **execution**: a user specifies the _what_
, _why_ and _how_ of an automatic data change and uses `dcmodifydb` to execute 
them on a `tbl` database table.

`dcmodidfydb` is optimized and restricted to database tables that can be accessed 
within R through `DBI`. It uses the `dbplyr` package to translate the 
data correction rules in R syntax into SQL syntax. The advantage of this 
approach is that all data correction is done within the database, which may be 
a requirement of your organisation or because the data table is simply too 
large to be held in memory. 
A disadvantage is that not all R statements can be translated into SQL statement,
so `dcmodifydb` is more restricted than `dcmodify` which can use the full R
potential. Nonetheless `dcmodifydb` may be sufficient and efficient for many
use cases.

For common error scenario's see `vignette("scenarios", package="dcmodifydb")`. 
For the supported syntax for specifying rules see `vignette("syntax", package="dcmodifydb")`.

## Installation

`dcmodifydb` can be installed with

```{r, eval=FALSE}
install.packages("dcmodifydb")
```
and loaded with:

```{r setup}
library(dcmodify)
library(dcmodifydb)
```

## Usage

`dcmodifydb` works on a database table, so we need a connection to a table within
a database. 

```{r}

```

