---
title: "dcmodifydb"
subtitle: "Documented reproducible data correction on a database"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dcmodifydb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The goal of dcmodifydb is to apply modification rules specified with `dcmodify` 
on a database table, allowing for documented, reproducable data cleaning adjustments
in a database.

`dcmodify` separates **intent** from **execution**: a user specifies the _what_
, _why_ and _how_ of an automatic data change and uses `dcmodifydb` to execute 
them on a `tbl` database table.

`dcmodidfydb` is optimized and restricted to database tables that can be accessed 
within R through `DBI`. It uses the `dbplyr` package to translate the 
data correction rules in R syntax into SQL syntax. The advantage of this 
approach is that all data correction is done within the database, which may be 
a requirement of your organisation or because the data table is simply too 
large to be held in memory. 
A disadvantage is that not all R statements can be translated into SQL statement,
so `dcmodifydb` is more restricted than `dcmodify` which can use the full R
potential. Nonetheless `dcmodifydb` may be sufficient and efficient for many
use cases.

## Installation

`dcmodifydb` can be installed with

```{r, eval=FALSE}
install.packages("dcmodifydb")
```
and loaded with:

```{r setup}
library(dcmodify)
library(dcmodifydb)
```

# Error scenario's

### Replacing common mistakes



```{r}
m <- modifier( if (year < 25) year = year + 2000)
```

```{r}
m <- modifier( if (year < 100){
  if (year > 25) {
    year = year + 1900
  } else {
    year = year + 2000
  }
})
```

### Replacing missing values

A value is not measured (`NA`), but can be deduced to be of a certain value, e.g. zero. Suppose a health questionair contains questions if a person smokes and if so how many cigarretes per day. 
Typically if the answer to the first question is "no", the second question is not asked an thus "unmeasured", but can be deduced to be 0.

```{r, results='asis', echo=FALSE}
d <- data.frame(smokes=c(TRUE, FALSE), cigarettes=c(10,NA))
knitr::kable(d)
```


```{r, eval = FALSE}
if (smokes ==  FALSE) cigarretes = 0
```

```{r, echo = FALSE, results='asis'}
m <- modifier(if (smokes ==  FALSE) {
  cigarretes = 0
})
dc <- modify(d, m)
knitr::kable(dc)
```


### Recoding


# SQL translation


# Syntax 

## `if` rule

```{r}
m <- modifier( if (age < 12) income = 0)
```

The following statements are equivalent. It is wise to choose a syntax
that is familiar to the persons specifying the correction rules. 

```{eval=FALSE}
if (age < 12) income = 0  # R noobs
if (age < 12) income <- 0 # a bit more R-y
income[age < 12] <- 0     # very R-y
```

### `else`

Each `if` rule may be followed with an `else` or `else if`

```{r}
m <- modifier(if (age > 67) {retired = TRUE} else {retired = FALSE})
```

The following statements are equivalent. It is wise to choose a syntax
that is familiar to the persons specifying the correction rules. 

```{eval=FALSE}
if (age > 67) {retired = TRUE} else {retired = FALSE} # R noobs
if (age > 67) {retired <- TRUE} else {retired <- FALSE} # R-y
retired <- if (age > 67) TRUE else FALSE # very R-y
retired <- age > 67 # very R-y
```


### multiple assignments

```{r}
m <- modifier( 
  if (age > 67) {
    retired = TRUE
    salary = 0 
  }
)
```

### `else if`

