---
title: "Correction scenarios"
subtitle: "Common errors and their corrections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Correction scenarios}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dcmodify)
library(dcmodifydb)
```

# Introduction

The goal of `dcmodifydb` is to apply modification rules specified with `dcmodify` 
on a database table, allowing for documented, reproducable data cleaning adjustments
in a database. 

This document provides examples and common error corrections scenario's for which 
`dcmodify` can be used. 

# Document rules

A good practice is to store the intent and description of each rule: this enables
future (co)workers on the dataset to evaluate the use of each rule. 
It is therefore recommended to store rules in a `yaml` format and document each
rule accordingly.

# Error scenario's

### Replacing common mistakes

```{r}
m <- modifier( if (year < 25) year = year + 2000)
```

```{r}
m <- modifier( if (year < 100){
  if (year > 25) {
    year = year + 1900
  } else {
    year = year + 2000
  }
})
```

### Replacing missing values

A value is not measured (`NA`), but can be deduced to be of a certain value, e.g. zero. Suppose a health questionair contains questions if a person smokes and if so how many cigarretes per day. 
Typically if the answer to the first question is "no", the second question is not asked an thus "unmeasured", but can be deduced to be 0.

```{r, results='asis', echo=FALSE}
d <- data.frame(smokes=c(TRUE, FALSE), cigarettes=c(10,NA))
knitr::kable(d)
```


```{r, eval = FALSE}
if (smokes ==  FALSE) cigarretes = 0
```

```{r, echo = FALSE, results='asis'}
m <- modifier(if (smokes ==  FALSE) {
  cigarretes = 0
})
dc <- modify(d, m)
knitr::kable(dc)
```


### Recoding


# SQL translation


# Syntax 

## `if` rule

```{r}
m <- modifier( if (age < 12) income = 0)
```

The following statements are equivalent. It is wise to choose a syntax
that is familiar to the persons specifying the correction rules. 

```{eval=FALSE}
if (age < 12) income = 0  # R noobs
if (age < 12) income <- 0 # a bit more R-y
income[age < 12] <- 0     # very R-y
```

### `else`

Each `if` rule may be followed with an `else` or `else if`

```{r}
m <- modifier(if (age > 67) {retired = TRUE} else {retired = FALSE})
```

The following statements are equivalent. It is wise to choose a syntax
that is familiar to the persons specifying the correction rules. 

```{eval=FALSE}
if (age > 67) {retired = TRUE} else {retired = FALSE} # R noobs
if (age > 67) {retired <- TRUE} else {retired <- FALSE} # R-y
retired <- if (age > 67) TRUE else FALSE # very R-y
retired <- age > 67 # very R-y
```


### multiple assignments

```{r}
m <- modifier( 
  if (age > 67) {
    retired = TRUE
    salary = 0 
  }
)
```

### `else if`

