---
title: "Modification syntax"
subtitle: "Supported rule syntax"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modification syntax}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dcmodify)
library(dcmodifydb)
```

# Introduction

The goal of dcmodifydb is to apply modification rules specified with `dcmodify` 
on a database table, allowing for documented, reproducable data cleaning adjustments
in a database.

This vignette describes the supported syntax that can be used with `dcmodify` and
`dcmodifydb`.

For common error scenario's see `vignette("scenarios", package="dcmodifydb")`


# Syntax 

## `if` rule

```{r}
m <- modifier( if (age < 12) income = 0)
```

The following statements are equivalent. It is wise to choose a syntax
that is familiar to the persons specifying the correction rules. 

```{eval=FALSE}
if (age < 12) income = 0  # R noobs
if (age < 12) income <- 0 # a bit more R-y
income[age < 12] <- 0     # very R-y
```

### `else`

Each `if` rule may be followed with an `else` or `else if`

```{r}
m <- modifier(if (age > 67) {retired = TRUE} else {retired = FALSE})
```

The following statements are equivalent. It is wise to choose a syntax
that is familiar to the persons specifying the correction rules. 

```{eval=FALSE}
if (age > 67) {retired = TRUE} else {retired = FALSE} # R noobs
if (age > 67) {retired <- TRUE} else {retired <- FALSE} # R-y
retired <- if (age > 67) TRUE else FALSE # very R-y
retired <- age > 67 # very R-y
```


### multiple assignments

```{r}
m <- modifier( 
  if (age > 67) {
    retired = TRUE
    salary = 0 
  }
)
```

### `else if`

